---
title: "Machine Learning Assignment"
author: "Bram van Zanten"
date: "Thursday, August 21, 2014"
output: html_document
---

1. Loading Data
===============
First, let's load the Caret library and the training (named main) and testing (named exam) data.

```{r}
library(caret)

main = read.csv("pml-training.csv")
exam  = read.csv("pml-testing.csv")
```

2. Preparing Data
=================
Now we split the training data (main) into training and testing dataset. We've kept the train part of the data small (p=0.10) for processing speed, ideally this should be a 40/60 split.
```{r}
base <- createDataPartition(y=main$classe, p=0.10, list=FALSE)
train <- main[base,]
test <- main[-base,]
```

3. Cleaning Data
=================
The next step is to proceed with some basic data cleaning. Variables with missing (NA & Empty) values are removed. 
We also remove some variables that would logically not be indicated to have predictive power for the question at hand (such as timestamp, username and row number).
```{r}
newtrain = train[,!sapply(train,function(x) any(is.na(x)))]
newtrain = newtrain[,!sapply(newtrain,function(x) any(x == ""))]
newtrain$X <- NULL
newtrain$raw_timestamp_part_1 <- NULL
newtrain$raw_timestamp_part_2 <- NULL
newtrain$num_window <- NULL 
newtrain$new_window <- NULL
newtrain$cvtd_timestamp <- NULL
newtrain$user_name <- NULL
```

4. Fitting Model
================
Next step is to setup the actual model. We're using a random forest algorithm.

```{r}
ctrl <- trainControl(method = "cv", number = 4, allowParallel = TRUE)
modelFit <- train(classe ~.,data=newtrain, method="rf", 
                  prof = TRUE, trControl = ctrl)
modelFit
```

As we can see from the accuracy measures, the indicicated predictive performance of the model is quite good. 

5. Cross-Validation
===================
To assess potential overfitting issues, we now cross-validate the model against the test data we set aside earlier. 

```{r}
pred <- predict(modelFit, test) 
test$predRight <- pred == 
test$classe 
table(pred, test$classe)
```


6. Preparing Answers
====================
Finally, we apply the model to the 'exam' (testing) data and use the provided function to write 20 text files.
```{r}

pred2 <- predict(modelFit, exam)
exam2 <- exam
exam2$classe <- pred2

pml_write_files = function(x) {
  n = length(x)
  for (i in 1:n) {
    filename = paste0("problem_id_", i, ".txt")
    write.table(x[i], file = filename, quote = FALSE, row.names = FALSE, 
                col.names = FALSE)
  }
}


answers <- exam2$classe

pml_write_files(answers)
answers
```